---
title: "STAT 545A homework 4"
author: "Tian Gao"
date: "2018/10/2"
output: github_document
---

# Bring rectangular data in
```{r}
## load gapminder
library(gapminder)
## load tidyverse
library(tidyverse)
```

# Data Reshaping Prompts (and relationship to aggregation)

# Activity #2

* Make a tibble with one row per year and columns for life expectancy for two or more countries.
* Use knitr::kable() to make this table look pretty in your rendered homework.
* Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

Filter the countries to Canada and China reshape the data, spread the country to columns, lifeExp as the values
```{r}

life_contrast<-gapminder %>%
# Filter the country to China and United States
  filter(country %in% c("China","Canada"))%>%
  select(year, country, lifeExp) %>%
# reshape the data, spread the country to columns, lifeExp as the values
  spread(country,lifeExp) %>%
  rename(Year = year)
knitr::kable(
  life_contrast
)
```

Then I'll do it for different countries, eg the countries of a continent
```{r}
life_contrast_asia <- gapminder %>%
  filter(continent == "Asia") %>%
  select(c("year", "country", "lifeExp")) %>%
  spread(country, lifeExp)

knitr::kable(
  life_contrast_asia
)

```

Then I'll draw a line plot of LifeExpectency of China and Canada
```{r}
# fix names with spaces
 life_contrast%>% 
  ggplot() +
  # make a line plot with points for China
  geom_line(aes(Year, China, color = "China")) +
  geom_point(aes(Year, China, color = "China")) +
  # make a line plot with points for Canada
  geom_line(aes(Year, Canada, color = "Canada")) +
  geom_point(aes(Year, Canada, color = "Canada")) +
  labs(y = "Life Expectency")
```

* Next I'll show the scatterplot that shows life expectancy for this two countries.
```{r}
life_contrast %>%
  ggplot(aes(China, Canada,color=factor(Year)))+
  geom_point()+
  xlim(40,85)+
  ylim(40,85)+
  coord_fixed()+
  geom_abline(intercept = 0, slope = 1)
  
```

# Join Prompts (join, merge, look up)

## Activity #1
* Problem: You have two data sources and you need info from both in one new data object.

* Solution: Perform a join, which borrows terminology from the database world, specifically SQL.

* Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:
* One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city.
* One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.

* country_codes in package gapminder is a good data frame for this situation. First, let's show the data in country_codes 
```{r}
country_codes %>% 
  head(30)%>%
  knitr::kable()
```

* full_join

```{r}
fullr <- full_join(gapminder,country_codes , by = "country") 
fullr %>% 
  head(30)%>%
  knitr::kable()
```

I'll check the dimension of result
```{r}
dim(fullr)
```

Then Let's change the pass in order to see what will happen.
```{r}
fulll <- full_join(gapminder,country_codes , by = "country") 
fulll %>% 
  head(30)%>%
  knitr::kable()

dim(fulll)
```

From the above experiment, we can see that full-join() with reversed pass in generated the same result. The reason is that full-join function shows all rows and columns from both x and y.

* inner_join()
```{r}
inner_join(gapminder, country_codes, by = "country") %>%
  head(30) %>%
  knitr::kable() 
```

```{r}
# shows the dimension of the result table
dim(inner_join(gapminder, country_codes, by = "country"))
```

Then I'll reverse the pass in parameters to see what will happen.
```{r}
inner_join(country_codes, gapminder, by = "country") %>% 
  head(30)%>%
  knitr::kable()
```

```{r}
dim(inner_join( country_codes, gapminder,by = "country"))
```

From the above experiment we can see that changing the order of passing in parameters does not change the result. Because the inner_join() only find the countries appeard in both data frames.

* semi_join()
```{r}
semi_join( gapminder, country_codes, by = "country") %>%
  head(30)%>%
  knitr::kable() 
```

I'll then change the order of parameters

```{r}
semi_join(country_codes, gapminder, by = "country") %>% 
  head(30)%>%
  knitr::kable() 
```

We can see that this time the results is different. Though semi_join() returns the overlapped countries, the results depend on the first parameter. 

* left_join()
```{r}
left_join(gapminder, country_codes, by = "country") %>%
  head(30)%>%
  knitr::kable() 
```

```{r}
left_join(country_codes, gapminder, by = "country")  %>%
  head(30) %>%
  knitr::kable() 
```

The results depends on the countries of the first parameter. if country_codes is the first parameter, then we can not find the countries only exist in gapminder. And for the countries only existed in country_codes, the value for that in gaominder is set to NA. 

* anti_join()
```{r}
anti_join(x = gapminder, y = country_codes, by = "country")
```

```{r}
anti_join(country_codes, gapminder, by = "country")  %>% 
  head(30)%>%
  knitr::kable()
```

anti_join() shows the rows in the first parameter that does not exist in the second one. Also the columns in the result is dependent on the first parameter. 




